---
title: "Final Project: ATP Matches & World Ranking Analysis"
author: "MSDS 597 Data Wrangling & Husbandry, Zishen Zhang"
date: "April 12, 2019"

output: 
  html_document: default
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```

## Package Lists:

```{r}

library(tidyverse)
library(data.table)
library(readr)
library(stringr)
library(rvest)
library(lubridate)
library(ggplot2)

```

## Some Prefix before Doing Analysis:

```{r}

atp <- read.csv(file = "ATP.csv", header = TRUE, stringsAsFactors = TRUE)
atp$tourney_name[atp$tourney_name == "Us Open"] <- "US Open"
atp$tourney_name[atp$tourney_name == "Australian Chps."] <- "Australian Open"
atp$tourney_date <- as.Date(atp$tourney_date, "%Y%m%d")
r <- with(atp, which(atp$tourney_name == "tourney_name", arr.ind = TRUE))
atp <- atp[-r, ]
atp <- atp[, -c(10, 16:17, 23, 25:27, 32:49)]
write.csv(atp, file = 'ATP Matches.csv')

```

## Figuring out the Number of Winning Games of Players in Their Career Lifes:

```{r}

tournament_level <- atp %>% 
  group_by(tourney_level) %>% 
  summarise(count = n()) %>% 
  arrange(tourney_level)
head(tournament_level)

winner_games <- atp %>% 
  group_by(winner_name) %>% 
  summarise(match_n = n()) %>% 
  arrange(desc(match_n)) %>% 
  select(winner_name, match_n)
head(winner_games, 10)

```

## Some Analysis Based on Grand Slam Results between 2003-2016：
## tourney_level == "G" means Grand Slams. 

```{r}

Grand_Slam_Matches <- atp %>%
  filter(substr(tourney_id, 1, 4) > "1967" & substr(tourney_id, 1, 4) < "2017" & tourney_level == "G")

Grand_Slam_Finals_2003_2016 <- Grand_Slam_Matches %>% 
  select(tourney_id, tourney_name, tourney_date, round, surface, winner_name, loser_name, score, minutes) %>%
  filter(substr(tourney_id, 1, 4) >= "2003" & round == "F") %>%
  arrange(desc(tourney_date, tourney_name))
Grand_Slam_Finals_2003_2016

ggplot(Grand_Slam_Finals_2003_2016, aes(winner_name, fill = as.factor(tourney_name))) +
  geom_bar(stat = "count", alpha = 1 ) + 
  labs(y = "Numbers of Titles", x = "Title Winners", title = "Grand Slams' Winners 2003 - 2016") +
  theme(legend.position = "right", axis.text.x = element_text(angle = 90, hjust = 0), plot.title = element_text(hjust = 0.5)) +
  guides(fill = guide_legend(title = "Tourney"))
ggplot(Grand_Slam_Finals_2003_2016, aes(winner_name, fill = as.factor(winner_name))) +
  geom_bar(alpha = 0.8, show.legend = FALSE) + 
  labs(y = "Titles", x = "Winners", title = "Grand Slams' Winners 2003 - 2016") +
  theme(axis.text.x = element_text(angle = 90, hjust = 0), plot.title = element_text(hjust = 0.5)) +
  coord_flip() +
  facet_wrap(~tourney_name)

```

## Some Analysis Based on ATP1000 Masters Series Results between 2003-2016：
## tourney_level == "M" means Masters Series.

```{r}

h <- with(atp, which(atp$tourney_name == "Hamburg Masters", arr.ind = TRUE))
atp <- atp[-h, ]

Masters_Matches <- atp %>%
  filter(substr(tourney_id, 1, 4) > "1967" & substr(tourney_id, 1, 4) < "2017" & tourney_level == "M")

Masters_Finals_2003_2016 <- Masters_Matches %>% 
  select(tourney_id, tourney_name, tourney_date, round, surface, winner_name, loser_name, score, minutes) %>%
  filter(substr(tourney_id, 1, 4) >= "2003" & round == "F") %>%
  arrange(desc(tourney_date, tourney_name))
Masters_Finals_2003_2016

ggplot(Masters_Finals_2003_2016, aes(winner_name, fill = as.factor(tourney_name))) +
  geom_bar(stat = "count", alpha = 1 ) + 
  labs(y = "Numbers of Titles", x = "Title Winners", title = "ATP Masters Series' Winners 2003 - 2016") +
  theme(legend.position = "right", axis.text.x = element_text(angle = 90, hjust = 0), plot.title = element_text(hjust = 0.5)) +
  guides(fill = guide_legend(title = "Tourney"))
ggplot(Masters_Finals_2003_2016, aes(winner_name, fill = as.factor(winner_name))) +
  geom_bar(alpha = 0.8, show.legend = FALSE) + 
  labs(y = "Titles", x = "Winners", title = "ATP Masters Series' Winners 2003 - 2016") +
  theme(axis.text.x = element_text(angle = 90, hjust = 0), plot.title = element_text(hjust = 0.5)) +
  coord_flip() +
  facet_wrap(~tourney_name)

```

## Extracting ATP World Ranking from ATP Official Website:

```{r}

ranking_2015 <- "https://www.atptour.com/en/rankings/singles?rankDate=2016-01-04&rankRange=1-10" %>% 
  read_html() %>% 
  html_table(fill = TRUE) %>%
  .[[1]]
ranking_2015 <- ranking_2015[, -c(2:3)]
ranking_2015

ranking_2016 <- "https://www.atptour.com/en/rankings/singles?rankDate=2017-01-02&rankRange=1-10" %>% 
  read_html() %>% 
  html_table(fill = TRUE) %>%
  .[[1]]
ranking_2016 <- ranking_2016[, -c(2:3)]
ranking_2016

```

## Unusual Patterns on Winner Games & World Ranking Anlysis

```{r}

Grand_Slam_Finals_2015_2016 <- Grand_Slam_Matches %>% 
  select(tourney_id, tourney_name, tourney_date, round, surface, winner_name, loser_name, score, minutes) %>%
  filter(substr(tourney_id, 1, 4) >= "2015" & round == "F") %>%
  arrange(desc(tourney_date, tourney_name))
Grand_Slam_Finals_2015_2016
ggplot(Grand_Slam_Finals_2015_2016, aes(winner_name, fill = as.factor(tourney_name))) +
  geom_bar(stat = "count", alpha = 1 ) + 
  labs(y = "Number of Finals", x = "Final Winner", title = "Grand Slams' Winners 2015 - 2016") +
  theme(legend.position = "right", axis.text.x = element_text(angle = 90, hjust = 0), plot.title = element_text(hjust = 0.5)) +
  guides(fill = guide_legend(title = "Tourney"))
ggplot(Grand_Slam_Finals_2015_2016, aes(winner_name, fill = as.factor(winner_name))) +
  geom_bar(alpha = 0.8, show.legend = FALSE) + 
  labs(y = "Finals", x = "Winners", title = "Grand Slams' Winners 2015 - 2016") +
  theme(axis.text.x = element_text(angle = 90, hjust = 0), plot.title = element_text(hjust = 0.5)) +
  coord_flip() +
  facet_wrap(~tourney_name)

Masters_Finals_2015_2016 <- Masters_Matches %>% 
  select(tourney_id, tourney_name, tourney_date, round, surface, winner_name, loser_name, score, minutes) %>%
  filter(substr(tourney_id, 1, 4) >= "2015" & round == "F") %>%
  arrange(desc(tourney_date, tourney_name))
Masters_Finals_2015_2016
ggplot(Masters_Finals_2015_2016, aes(winner_name, fill = as.factor(tourney_name))) +
  geom_bar(stat = "count", alpha = 1 ) + 
  labs(y = "Number of Finals", x = "Final Winner", title = "ATP Masters Series' Winners 2015 - 2016") +
  theme(legend.position = "right", axis.text.x = element_text(angle = 90, hjust = 0), plot.title = element_text(hjust = 0.5)) +
  guides(fill = guide_legend(title = "Tourney"))
ggplot(Masters_Finals_2015_2016, aes(winner_name, fill = as.factor(winner_name))) +
  geom_bar(alpha = 0.8, show.legend = FALSE) + 
  labs(y = "Finals", x = "Winners", title = "ATP Masters Series' Winners 2015 - 2016") +
  theme(axis.text.x = element_text(angle = 90, hjust = 0), plot.title = element_text(hjust = 0.5)) +
  coord_flip() +
  facet_wrap(~tourney_name)

```

```